Implicit assumptions and their impact on photometric redshift PDF performance for LSST

### 2019-02-28 paper/tex/analysis.tex

The exact implementation of the stacked estimator $\hat{N}^{H}(z)$ should depend on the parametrization of the \pzpdf s, which may differ across codes and can affect the precision of the estimator \citep{Malz:qp}; even considering a single method under the same parametrization, say a piecewise constant function over bins or a set of samples from the posterior, an estimator using $2N$ bins or samples will trivially be more precise than an estimator using $N$ bins or samples.
%\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}

In order to minimize the effects of the choice of parameterization, we asked those running all twelve codes to output \pzpdf s parameterized with $\approx 200$ piecewise constant bins spanning $0<z<2$.
The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of \pzpdf s for the final \lsst\ Project tables.\footnote{See, e.~g.~the \lsst\ Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
All the \pzpdf\ catalogs are processed using the \qp\ software package \citep{Malz:qp}\footnote{available at: \url{http://github.com/aimalz/qp/}} for manipulating and calculating metrics of univariate PDFs.
We will discuss the choice of \pzpdf\ parameterization further in Section~\ref{sec:results}.

@@

\subsubsection{Root-mean-square error (RMSE)}
\label{sec:rmse}
%(Alex Malz)

We employ the familiar root-mean-square error
\begin{equation}
\mathrm{RMSE} \equiv \sqrt{\int_{-\infty}^{\infty}\left(\hat{f}(z)-f'(z)\right)^{2}\,dz} .
\end{equation}
%may be used for two purposes in this analysis.
%We calculate it between the true redshift distribution $N'(z)$ and each estimator $\hat{N}(z)$ (derived by Eq.~\ref{eq:stacked}) as well as between the ideal QQ and PIT curves from each method studied and the ideal curves.
Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution.%, and it can be used to quantify the otherwise qualitative metrics.

@@

We additionaly calculate the first three moments of the estimated redshift distribution $\hat{N}^{H}(z)$ for each code and compare them to the moments of the true redshift distribution $N'(z)$.
The $m^{\mathrm{th}}$ moment of a distribution is defined as
\begin{equation}
\langle z^{m}\rangle \equiv \int_{-\infty}^{\infty}z^{m}N(z)dz .
\end{equation}
Here, we use the moments of the stacked estimator of the redshift distribution function as the basis for a metric.
The closer the moments of $\hat{N}(z)$ for a \pzpdf\ method are to the moments of the true redshift distribution function $N'(z)$, the better the \pzpdf\ method.

>>>>

The exact implementation of the stacked estimator $\hat{N}^{H}(z)$ should depend on the parametrization of the \pzpdf s, which may differ across codes and can affect the precision of the estimator \citep{Malz:qp}; even considering a single method under the same parametrization, say a piecewise constant function over bins or a set of samples from the posterior, an estimator using $2N$ bins or samples will trivially be more precise than an estimator using $N$ bins or samples.
%\red{Note: Let's address how the gridding (endpoints and spacing) affects the metrics, from the email correspondence Re: [LSST-DESC-MEMBERS] Summary of Nov 29 PZ Telecon. (AIM)}
In order to minimize the effects of the choice of parameterization, we asked those running all twelve codes to output \pzpdf s parameterized with $200$ piecewise constant bins spanning $0<z<2$.
The piecewise constant format is chosen because of its established presence in the literature, and the choice of 200 bins was motivated by the approximate number of columns expected to be available for storage of \pzpdf s for the final \lsst\ Project tables.\footnote{See, e.~g.~the \lsst\ Data Products Definition Document, available at: \url{https://ls.st/dpdd}}
All the \pzpdf\ catalogs are processed using the \qp\ software package \citep{Malz:qp}\footnote{available at: \url{http://github.com/aimalz/qp/}} for manipulating and calculating metrics of univariate PDFs.
We will discuss the choice of \pzpdf\ parameterization further in Section~\ref{sec:results}.

@@

%\subsubsection{Root-mean-square error (RMSE)}
%\label{sec:rmse}
%%(Alex Malz)%
%
%We employ the familiar root-mean-square error
%\begin{equation}
%\mathrm{RMSE} \equiv \sqrt{\int_{-\infty}^{\infty}\left(\hat{f}(z)-f'(z)\right)^{2}\,dz} .
%\end{equation}
%%may be used for two purposes in this analysis.
%%We calculate it between the true redshift distribution $N'(z)$ and each estimator $\hat{N}(z)$ (derived by Eq.~\ref{eq:stacked}) as well as between the ideal QQ and PIT curves from each method studied and the ideal curves.
%Though this metric does not account for the fact that the redshift distribution function is, in fact, a probability distribution, it can still be interpreted as a measure of the integrated difference between the estimated distribution and the true distribution.%, and it can be used to quantify the otherwise qualitative metrics.
%

@@

We additionaly calculate the first three moments of the estimated redshift distribution $\hat{N}^{H}(z)$ for each code and compare them to the moments of the true redshift distribution $N'(z)$.
The $m^{\mathrm{th}}$ moment of a distribution is defined as
\begin{equation}
\langle z^{m}\rangle \equiv \int_{-\infty}^{\infty}z^{m}N(z)dz .
\end{equation}
Here, we use the moments of the stacked estimator of the redshift distribution function as the basis for a metric.
The closer the moments of $\hat{N}(z)$ for a \pzpdf\ method are to the moments of the true redshift distribution function $N'(z)$, the better the \pzpdf\ method is a estimating the overall redshift distribution ``shape''.

### 2019-02-28 paper/tex/data.tex

\subsubsection{Selection of training and test sets}
\label{sec:buzztraining}
The total catalogue covered $400$ square degrees and contained $238$ million galaxies to an apparent magnitude limit of $r\!=\!29$ and spanning the redshift range $0\!<\!z\!\leq\!8.7$.  In order for statistical errors not to dominate, we need less than one million galaxies in our sample.  Several studies claim that only a few tens of thousands of spectra are necessary to calibrate photo-z surveys to Stage IV requirements (e.~g.~\citet{Bernstein:10}, \citet{Masters:2017}).  Therefore, we aim for a final number of training galaxies between $3\times 10^{4}$ and $5\times 10^{4}$ in our sample.
%This catalogue contained two orders of magnitude more galaxies than were necessary to determine statistics for this study,
In order to reduce our sample to a reasonable size, we limit our dataset to a subset of $\sim\!16.8$ square degrees selected from five separate spatial regions of the simulation. Systematic problems with galaxy colors above $z\!>\!2$ were observed, so the catalogue was limited to include only galaxies in the redshift range $0\!<\!z\!\leq\!2.0$. A random subset of the the remaining galaxies was chosen, and placed at random into either a ``training'' set ($10$ per cent of the sample), for which the galaxies true redshifts will be supplied, or a ``test'' set (the remaining $90$ per cent of the sample), for which each code will need to predict a redshift PDF for each galaxy.  
%The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
Finally, we restrict our analysis to a sample with an apparent magnitude limite $i<25.3$, which give a signal-to-noise $\sim\,30$ for most galaxies, a cut often referred to as the expected ``LSST Gold Sample''.  This magnitude cut results in a training set with $44\,404$ galaxies and a test set containing $399\,356$ galaxies.  All subsequent results will evaluate this ``gold sample'' test set.  In order to blind results, initially redshifts were not revealed for the ``test'' set, and were only supplied for the training sample galaxies.  This prevented code runners from tweaking results and over-fitting to the specific test set.

@@

\subsubsection{Limitations}
\label{sec:buzzlimitations}
For our initial investigation of photometric redshift codes, we begin with a data set that is somewhat idealized, and does not contain all of the complicating factors present in real data.  In several cases, the simplification is done with a purpose, with potentially confounding effects excluded in order to better isolate the differences between current-generation photo-$z$ codes, and their causes.  We list several of the simulations limitations in this section.
As the simulation is catalogue-based, no image level effects, such as photometric measurement effects, object blending, contamination from sky background (Zodiacal light, scattered light, etc...), lensing magnification, or Galactic reddening are included.  No stars are included in the catalogue, nor are the effects of AGN.
As all SEDs are constructed from only five basis templates, properties of the galaxy population will be restricted to follow linear combinations of the characteristics of the five basis templates, so certain non-linear features, for example the full range of emission line fluxes relative to the continuum, will not be included in the model galaxy population.  Moreover, the linear combinations of templates are modeled on the $\sim5\times 10^{5}$ SDSS galaxies discussed in Section~\ref{sec:buzzard}, and thus only galaxies that resemble those spectroscopically observed by the SDSS will be included in the sample.  No additional dust reddening intrinsic to the host galaxy is included, the only approximation of dust extinction comes in the form of dust encoded in the five basis SEDs via the training set used to create the basis templates.  Simple linear combinations of these basis templates will, once again, not explore the full range of realistic dust extinction observed in galaxy populations.
While these idealized conditions limit the realism of our galaxy population, some are also by design.  We aim to test the photo-z codes at a very basic level, and a simplified model assures that differences in results seen between the codes are due to fundamental differences in their underlying assumptions and implementation details, rather than more nuanced implementation details.

>>>>

\subsubsection{Selection of training and test sets}
\label{sec:buzztraining}
The total initial catalogue covered $400$ square degrees and contained $238$ million galaxies to an apparent magnitude limit of $r\!=\!29$ and spanning the redshift range $0\!<\!z\!\leq\!8.7$.  In order for statistical errors not to dominate, we need less than one million galaxies in our sample.  Several studies claim that only a few tens of thousands of spectra are necessary to calibrate photo-z surveys to Stage IV requirements (e.~g.~\citet{Bernstein:10}, \citet{Masters:2017}).  Therefore, we aim for a final number of training galaxies between $3\times 10^{4}$ and $5\times 10^{4}$ in our sample.
%This catalogue contained two orders of magnitude more galaxies than were necessary to determine statistics for this study,
In order to reduce our sample to a reasonable size, we limit our dataset to a subset of $\sim\!16.8$ square degrees selected from five separate spatial regions of the simulation. Systematic problems with galaxy colors above $z\!>\!2$ were observed, so the catalogue was limited to include only galaxies in the redshift range $0\!<\!z\!\leq\!2.0$. A random subset of the the remaining galaxies was chosen, and placed at random into either a ``training'' set ($10$ per cent of the sample), for which the galaxies true redshifts will be supplied, or a ``test'' set (the remaining $90$ per cent of the sample), for which each code will need to predict a redshift PDF for each galaxy.  
%The resulting catalogues contain $111\,171$ training galaxies and $1\,000\,883$ test galaxies.
Finally, we restrict our analysis to a sample with an apparent i-band magnitude limit $i<25.3$, which give a signal-to-noise $\sim\,30$ for most galaxies, a cut often referred to as the expected ``LSST Gold Sample''.  This magnitude cut results in a training set with $44\,404$ galaxies and a test set containing $399\,356$ galaxies.  All subsequent results will evaluate this ``gold sample'' test set.  In order to blind results, initially redshifts were not revealed for the ``test'' set, and were only supplied for the training sample galaxies.  This prevented code runners from tweaking results and fitting to the specific test set.

@@

\subsubsection{Limitations}
\label{sec:buzzlimitations}
For our initial investigation of photometric redshift codes, we begin with a data set that is somewhat idealized, and does not contain all of the complicating factors present in real data.  In several cases, the simplification is done with a purpose, with potentially confounding effects excluded in order to better isolate the differences between current-generation photo-$z$ codes and their causes.  We list several of the simulation limitations in this section.
As the simulation is catalogue-based, no image level effects, such as photometric measurement effects, object blending, contamination from sky background (Zodiacal light, scattered light, etc...), lensing magnification, or Galactic reddening are included.  No stars are included in the catalogue, nor are the effects of AGN.
As all SEDs are constructed from only five basis templates, properties of the galaxy population will be restricted to follow linear combinations of the characteristics of the five basis templates, so certain non-linear features, for example the full range of emission line fluxes relative to the continuum, will not be included in the model galaxy population.  Moreover, the linear combinations of templates are modeled on the $\sim5\times 10^{5}$ SDSS galaxies discussed in Section~\ref{sec:buzzard}, and thus only galaxies that resemble those spectroscopically observed by the SDSS will be included in the sample.  No additional dust reddening intrinsic to the host galaxy is included, the only approximation of dust extinction comes in the form of dust encoded in the five basis SEDs via the training set used to create the basis templates.  Simple linear combinations of these basis templates will, once again, not explore the full range of realistic dust extinction observed in galaxy populations.
While these idealized conditions limit the realism of our galaxy population, some are also by design.  We aim to test the photo-z codes at a very basic level, and a simplified model assures that differences in results seen between the codes are due to fundamental differences in their underlying assumptions and implementation details, rather than more nuanced properties.

### 2019-02-28 paper/tex/methods.tex

Here we outline the photo-$z$ PDF codes tested in this study. In total, eleven distinct codes are tested.  This sample is not comprehensive, codes were chosen based on the expertise available within the group; however, those chosen do cover a broad range of the current-generation methods used in the field.  Both template-based and machine learning approaches are included and each are described separately in Secs.~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively. The list of codes are summarized in Table.~\ref{tab:list_of_codes}.  All code runners were asked to output redshift posterior estimates on 200 linear-spaced bins between redshifts 0 and 2.

@@

The nearest-neighbours colour-matching photometric redshift estimator (\textsc{CMNN}) is presented in \citet[][hereafter G18]{Graham:17}. This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage. For each galaxy in the test set we identify a colour-matched subset of training galaxies.  This subset is identified by first calculating the Mahalanobis distance $D_M$ in colour-space between the test galaxy and all training-set galaxies:
\begin{equation} 
D_{\rm M} = \sum^{N_{\rm colours}}\frac{(c_{\rm train}-c_{\rm test})^2}{(\delta c_{\rm test})^2}
\end{equation}
where c colour, $\delta c_{\rm test}$ is the measurement error on the colour, and $N_{\rm colours}$ is the total number of colors (i.e., in our case $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$). Then, we choose a threshold value for $D_M$ that defines the colour match set based on a set value of the percent point function (PPF): for example, for $N_{\rm dof}=5$, choosing a vPPF$=95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$ is the number of degrees of freedom, in this case the number of colours). If a galaxy had a non-detection in a band, that band was dropped and $N_{\rm dof}$ was reduced by one in the colour-matching space.  For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this colour-matched subset of training galaxies.

We have applied the nearest-neighbours colour-matching photometric redshift estimator described in G18 to the simulated data. Compared to its application in G18, there are some minor differences in the application of this estimator to the Buzzard catalogue. First, we do not impose non-detections on galaxies with a magnitude fainter than the expected LSST 10-year limiting magnitude or bright enough to saturate with LSST: {\it all} of the photometry for all the galaxies in the test and training sets are used for this experiment. Second, as in G18 we do apply an initial cut in colour to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colours and magnitudes. Third, we set different parameters for the identification of the colour-matched subset of training galaxies and the selection of a photometric redshift estimate. In G18 we used a percent point function (PPF) value of 0.68 to identify the colour-matched subset of training galaxies and used the redshift of nearest neighbour in colour-space as the photo-$z$ estimate. These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return an accurate $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in colour-space. Since an accurate estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator. We continue to use a percent point function of ${\rm PPF} = 0.95$ to generate the subset of colour-matched training galaxies, but weight them by the inverse of their Mahalanobis distance. This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in colour-space. We then use the weights to create the $p(z)$ instead of having the redshift of each colour-matched training-set galaxy count equally. To obtain a robust estimate of the $p(z)$ for galaxies with a small number of colour-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in colour-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:
\begin{equation}
\sigma = \sigma_{\rm train} \sqrt{({\rm PPF}_{20}/0.95)^2 -1}
\end{equation}
\noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it. Overall, these three changes will yield less precise photo-$z$ estimates compared to those presented in G18, but they will all have significantly more accurate estimates of the $p(z)$, particularly for the brightest test galaxies. This is sufficient for this work because, as described in G18, the goal of the \textsc{CMNN} photo-$z$ estimator was never to provide the ``best'' (or even competitive) estimates in the first place, given its reliance on a deep training set, but rather to provide a means for direct comparisons between LSST photometric quality and photo-$z$ estimates. With this work we show how the input parameters should be set in order to return accurate $p(z)$ estimates in addition to point value estimates.

@@

\textsc{GPz}\footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, with the added feature of being able to produce input-dependent variance estimations (heteroscedastic noise). The model assumes that the probability of the output $y$, the redshift, given the input $x$, the photometry, is $p(y|x)=\mathcal{N}\left(y|\mu(x),\sigma(x)^{2}\right)$. The mean function, $\mu(x)$, and the variance function $\sigma(x)^{2}$ are both linear combinations of basis functions that take the following form:
\begin{equation}
f(x)=\sum_{i=1}^{m}\phi_{i}(x)w_{i},
\end{equation}
where $\left\{\phi_{i}(x)\right\}_{i=1}^{m}$ and $\left\{w_{i}\right\}_{i=1}^{m}$ are sets of $m$ basis functions and their associated weights respectively. Basis function models (BFM), for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e. there exist a function of that form that can approximate any function, with mild assumptions, to any desired degree of accuracy. The details on how to learn the parameters of the model and the hyper-parameters of the basis functions are described in \citet{Almosallam:15b}.

@@

\textsc{METAPhoR} includes data modules for pre-processing, photo-$z$ estimation, and PDF estimation, and post-processing.  The pre-processing includes a model for perturbation of the photometry that is employed in calculating the PDF of the photo-$z$ estimation errors.  The photometric perturbation is defined as: $m_{ij} = m_{ij} + \alpha_{i}F_{ij}*u_{\mu=0,\sigma=1}$, where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.

>>>>

Here we outline the photo-$z$ PDF codes tested in this study. In total, twelve distinct codes are tested.  This sample is not comprehensive, codes were chosen based on the expertise available within the group; however, those chosen do cover a broad range of the current-generation methods used in the field.  Both template-based and machine learning approaches are included and each are described separately in Secs.~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively. The list of codes are summarized in Table.~\ref{tab:list_of_codes}.  All code runners were asked to output redshift posterior estimates on 200 linear-spaced bins between redshifts 0 and 2.

@@

The nearest-neighbours colour-matching photometric redshift estimator (\textsc{CMNN}) is presented in \citet[][hereafter G18]{Graham:17}. This method uses a training set of galaxies with known redshifts that has equivalent or better photometry as the test set in terms of quality and filter coverage. For each galaxy in the test set we identify a colour-matched subset of training galaxies.  This subset is identified by first calculating the Mahalanobis distance $D_M$ in colour-space between the test galaxy and all training-set galaxies:
\begin{equation} 
D_{\rm M} = \sum^{N_{\rm colours}}\frac{(c_{\rm train}-c_{\rm test})^2}{(\delta c_{\rm test})^2}
\end{equation}
where c colour, $\delta c_{\rm test}$ is the measurement error on the colour, and $N_{\rm colours}$ is the total number of colors (i.e., in our case $u$-$g$, $g$-$r$, $r$-$i$, $i$-$z$, and $z$-$y$). Then, we choose a threshold value for $D_M$ that defines the colour match set based on a set value of the percent point function (PPF): for example, for $N_{\rm dof}=5$, choosing a vPPF$=0.95$, $95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$ (where $N_{\rm dof}$ is the number of degrees of freedom, in this case the number of colours). If a galaxy had a non-detection in a band, that band was dropped and $N_{\rm dof}$ was reduced by one in the colour-matching space.  For a given test galaxy, the $p(z)$ is the normalized distribution of the true catalogue redshifts of this colour-matched subset of training galaxies.

We have applied the nearest-neighbours colour-matching photometric redshift estimator described in G18 to the simulated data. Compared to its application in G18, there are some minor differences in the application of this estimator to the Buzzard catalogue. First, we do not impose non-detections on galaxies with a magnitude fainter than the expected LSST 10-year limiting magnitude or bright enough to saturate with LSST: {\it all} of the photometry for all the galaxies in the test and training sets are used for this experiment. Second, as in G18 we do apply an initial cut in colour to the training set before calculating the Mahalanobis distance in order to accelerate processing, and also use a magnitude pseudo-prior to improve photo-$z$ estimates, but for both we have used different cut-off values that are appropriate for the Buzzard galaxies' colours and magnitudes. Third, we set different parameters for the identification of the colour-matched subset of training galaxies and the selection of a photometric redshift estimate. In G18 we used a PPF value of 0.68 to identify the colour-matched subset of training galaxies and used the redshift of nearest neighbour in colour-space as the photo-$z$ estimate. These choices work well when the desire is to obtain accurate photo-$z$ estimates for most test-set galaxies, but does not return an accurate $p(z)$ in all cases -- especially for galaxies that are bright and/or have few matches in colour-space. Since an accurate estimate of $p(z)$ is desired for this work we make several changes to our implementation of the \textsc{CMNN} photo-$z$ estimator. We continue to use a ${\rm PPF} = 0.95$ to generate the subset of colour-matched training galaxies, but weight them by the inverse of their Mahalanobis distance. This weighting maintains some of the accuracy that was previously achieved by simply using the nearest neighbour in colour-space. We then use the weights to create the $p(z)$ instead of having the redshift of each colour-matched training-set galaxy count equally. To obtain a robust estimate of the $p(z)$ for galaxies with a small number of colour-matched training set galaxies, when this number is less than $20$ the nearest $20$ neighbours in colour-space are used instead, and we convolve the $p(z)$ with a Gaussian with a standard deviation of:
\begin{equation}
\sigma = \sigma_{\rm train} \sqrt{({\rm PPF}_{20}/0.95)^2 -1}
\end{equation}
\noindent to appropriately broaden it so that the $p(z)$ for these test galaxies represents the enlarged PPF value associated with it. Overall, these three changes will yield less precise photo-$z$ estimates compared to those presented in G18, but they will all have significantly more accurate estimates of the $p(z)$, particularly for the brightest test galaxies. This is sufficient for this work because, as described in G18, the goal of the \textsc{CMNN} photo-$z$ estimator was never to provide the ``best'' (or even competitive) estimates in the first place, given its reliance on a deep training set, but rather to provide a means for direct comparisons between LSST photometric quality and photo-$z$ estimates. With this work we show how the input parameters should be set in order to return accurate $p(z)$ estimates in addition to point value estimates.

@@

\textsc{GPz}\footnote{\url{https://github.com/OxfordML/GPz}} \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, with the added feature of being able to produce input-dependent variance estimations (heteroscedastic noise). The model assumes that the probability of the output $y$, the redshift, given the input $x$, the photometry, is $p(z|x)=\mathcal{N}\left(z|\mu(x),\sigma(x)^{2}\right)$. The mean function, $\mu(x)$, and the variance function $\sigma(x)^{2}$ are both linear combinations of basis functions that take the following form:
\begin{equation}
f(x)=\sum_{i=1}^{m}\phi_{i}(x)w_{i},
\end{equation}
where $\left\{\phi_{i}(x)\right\}_{i=1}^{m}$ and $\left\{w_{i}\right\}_{i=1}^{m}$ are sets of $m$ basis functions and their associated weights respectively. Basis function models (BFM), for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e. there exist a function of that form that can approximate any function, with mild assumptions, to any desired degree of accuracy. The details on how to learn the parameters of the model and the hyper-parameters of the basis functions are described in \citet{Almosallam:15b}.

@@

\textsc{METAPhoR} includes data modules for pre-processing, photo-$z$ estimation, and PDF estimation, and post-processing.  The pre-processing includes a model for perturbation of the photometry that is employed in calculating the PDF of the photo-$z$ estimation errors.  The photometric perturbation is defined as:
\begin{equation}
 m_{ij} = m_{ij} + \alpha_{i}F_{ij}u_{\mu=0,\sigma=1}
\end{equation}
\noindent  where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.
